Copulas and their potential for ecology

### 2020-03-02 Paper.Rmd
legend color modified
small changes on correcting of AER proofs

Plots using normalized ranks were proposed in the Introduction as being conceptually
similar to copulas; we explain the connection. As stated above, the copula associated
with a random variable $(X,Y)$ is the cdf of the random variable
$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$ from $(X,Y)$, let $\hat{F}_X$ and
$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
$x$ and $y$ and jump by $1/n$
at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
the set $\{x_i : i=1,\ldots,n\}$, divided by $n$, 
and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
$\{y_i : i=1,\dots,n\}$, divided
by $n$. For large $n$, $\hat{F}_X(x_i)$ and $\hat{F}_Y(y_i)$ 
approximate the cdfs $F_X$ and $F_Y$, respectively. 
But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
which is the random variable associated with the copula of $(X,Y)$. The scatterplot 
of $v_i$ versus $u_i$ can be used to infer aspects of copula 
structure (see section \ref{Methods} below).

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
\caption{Upper panels show raw data plots for (A) $\log_{10}$(soil C) and $\log_{10}$(soil N) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
for birds, (C) the same for mammals, and (D) above-ground plant biomass vs. Shannon's index (\textit{H}) from Cedar Creek. Bottom panels show the corresponding normalized rank plots. See section \ref{Data} for units used in upper panels. \label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}

>>>>

Plots using normalized ranks were proposed in the Introduction as being conceptually
similar to copulas; we explain the connection. As stated above, the copula associated
with a random variable $(X,Y)$ is the cdf of the random variable
$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$ from $(X,Y)$, let $\hat{F}_X$ and
$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
$x$ and $y$ and jump by $1/n$
at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
the set $\{x_i : i=1,\ldots,n\}$, divided by $n$, 
and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
$\{y_i : i=1,\dots,n\}$, divided
by $n$. For large $n$, $\hat{F}_X(x_i)$ and $\hat{F}_Y(y_i)$ 
approximate the cdfs $F_X$ and $F_Y$, respectively. 
But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
$(n+1)/n$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
$(n+1)/n$. So the normalized rank in turn approximates the empirical cdf and therefore 
the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
which is the random variable associated with the copula of $(X,Y)$. The scatterplot 
of $v_i$ versus $u_i$ can be used to infer aspects of copula 
structure (see section \ref{Methods} below).

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
\caption{Upper panels show raw data plots for (A) $\log_{10}$(soil N) vs. $\log_{10}$(soil C) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
for birds, (C) the same for mammals, and (D) above-ground plant biomass vs. Shannon's index from Cedar Creek. Bottom panels show the corresponding normalized rank plots. See section \ref{Data} for units used in upper panels. \label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}
